---
title: "Temperature Modeling Restults and Plots"
author: "Anna Boser"
date: "1/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```


```{r getdata, echo=FALSE}
#first read in the dataset
Comp_temp <- readRDS(here::here("temperature_modeling", 
                              "data", 
                              "processed_data", 
                              "merged_df.RData"))
```


### Figure 3: LST vs. Air temp scatterplot
```{r lst_scatter, echo=FALSE, warning=FALSE}
ggplot(Comp_temp, aes(x = ECOSTRESS, y =Air_Temp)) + 
  geom_point(aes(alpha = .2)) +
  # stat_function(fun = function(x){6.906e-01 + 1.432e+00*x -2.213e-02*x^2 + 1.473e-04*x^3}, aes(color = "Cubic polynomial"), show.legend = TRUE) + 
  stat_function(fun = function(x){x}, aes(color = "One to one"), show.legend = TRUE) + 
  labs(title = "Land surface temperature vs. air temperature in the San Joaquin Valley", 
       subtitle = "June - September, 2018 - 2020", 
       caption = "Air temperature values from CIMIS weather stations.
       Land surface temperature from ECOSTRESS") + 
  ylab("Air Temperature (C)") + 
  xlab("Land Surface Temperature (C)") + 
  ylim(min = 5, max = 45) + 
  xlim(min = 5, max = 60) + 
  guides(alpha = FALSE) +
  theme(legend.title = element_blank())
```

### Figure 4: breakdown vs vegetation
```{r vegetation, warning = FALSE, echo = FALSE}
# ggplot(Comp_temp, aes(x =  ECOSTRESS, y =Air_Temp)) + 
#   geom_point(aes(color = V, alpha = .2)) +
#   # stat_function(fun = function(x){6.906e-01 + 1.432e+00*x -2.213e-02*x^2 + 1.473e-04*x^3}, show.legend = TRUE) + 
#   stat_function(fun = function(x){x}, show.legend = TRUE) + 
#   labs(title = "Land surface temperature vs. air temperature by vegetation fraction", 
#        subtitle = "June - September, 2018 - 2020, San Joaquin Valley", 
#        caption = "Air temperature values from CIMIS weather stations.
#        Land surface temperature from ECOSTRESS") + 
#   ylab("Air Temperature (C)") + 
#   xlab("Land Surface Temperature (C)") + 
#   ylim(min = 0, max = 45) + 
#   xlim(min = 0, max = 60) + 
#   scale_colour_gradient(low = "tan", high = "darkgreen") + 
#   labs(color = "Vegetation fraction") + 
#   # theme_dark() + 
#   guides(alpha = FALSE)

ggplot(Comp_temp, aes(x =  V, y =Air_Temp)) + 
    geom_point(aes(color = ECOSTRESS, alpha = .2)) +
  # stat_function(fun = function(x){6.906e-01 + 1.432e+00*x -2.213e-02*x^2 + 1.473e-04*x^3}, show.legend = TRUE) + 
  geom_smooth(method = "lm") + 
  labs(title = "Fractional vegetation vs. air temperature", 
       subtitle = "June - September, 2018 - 2020, San Joaquin Valley") + 
  ylab("Air Temperature (C)") + 
  xlab("Fractional vegetation") + 
  ylim(min = 0, max = 45) + 
  scale_colour_gradient(low = "green3", high = "red") + 
  labs(color = "Land Surface Temperature") + 
  # theme_dark() + 
  guides(alpha = FALSE)

```

What model should we choose? 

```{r choose model, warning=FALSE, echo=FALSE, include = FALSE}
# models to choose from: 
# simple linear 
# quardrtic
# cubic
# with and without veg
# what about veg interactions? 

lm1 <- lm(Air_Temp ~ ECOSTRESS, data = Comp_temp)
lm2 <- lm(Air_Temp ~ ECOSTRESS + I(ECOSTRESS^2), data = Comp_temp)
lm3 <- lm(Air_Temp ~ ECOSTRESS + I(ECOSTRESS^2) + I(ECOSTRESS^3), data = Comp_temp)
lm4 <- lm(Air_Temp ~ ECOSTRESS + I(ECOSTRESS^2) + V, data = Comp_temp)

ggplot(Comp_temp, aes(y =lm1$resid, x = ECOSTRESS)) + 
    geom_point(aes(color = V, alpha = .2)) +
  geom_smooth(method = "lm") + 
  labs(title = "Fractional vegetation vs. air temperature", 
       subtitle = "June - September, 2018 - 2020, San Joaquin Valley") + 
  scale_colour_gradient(low = "tan", high = "green3") + 
  labs(color = "Fv") + 
  guides(alpha = FALSE)

ggplot(Comp_temp, aes(y =lm2$resid, x = ECOSTRESS)) + 
    geom_point(aes(color = V, alpha = .2)) +
  geom_smooth(method = "lm") + 
  labs(title = "Fractional vegetation vs. air temperature", 
       subtitle = "June - September, 2018 - 2020, San Joaquin Valley") + 
  scale_colour_gradient(low = "tan", high = "green3") + 
  labs(color = "Fv") + 
  guides(alpha = FALSE)

ggplot(Comp_temp, aes(y =lm3$resid, x = ECOSTRESS)) + 
    geom_point(aes(color = V, alpha = .2)) +
  geom_smooth(method = "lm") + 
  labs(title = "Fractional vegetation vs. air temperature", 
       subtitle = "June - September, 2018 - 2020, San Joaquin Valley") + 
  scale_colour_gradient(low = "tan", high = "green3") + 
  labs(color = "Fv") + 
  guides(alpha = FALSE)

ggplot(Comp_temp, aes(y =lm4$resid, x = ECOSTRESS)) + 
    geom_point(aes(color = V, alpha = .2)) +
  geom_smooth(method = "lm") + 
  labs(title = "Fractional vegetation vs. air temperature", 
       subtitle = "June - September, 2018 - 2020, San Joaquin Valley") + 
  scale_colour_gradient(low = "tan", high = "green3") + 
  labs(color = "Fv") + 
  guides(alpha = FALSE)
```

When choosing between the linear, quadratic, and cubic, though they're all statistically significant, we choose quadratic because:

 * the linear residuals are not homoscedastic
 * the quadratic residuals are homoscedastic and the cubic only increases the Adjusted $R^2$ by .001 (0.8474 to 0.8488). 

We then do not include vegetation in the quadratic because: 

 * It is not statistically significant and the adjusted $R^2$ actually decreases by .001

### Figure 5: Corrected relationship w/ 1:1 scatter

```{r 1to1_scatter, warning = FALSE, echo = FALSE}
#chosen model
final <- lm2

#fit values
Comp_temp$air_temp_modeled <- final$fitted.values

ggplot(Comp_temp, aes(y = Air_Temp, x = air_temp_modeled)) + 
  geom_point(aes(alpha = .2)) +
  stat_function(fun = function(x){x}, aes(color = "One to one"), show.legend = TRUE) + 
  labs(title = "Modeled vs. Observed Air Temperature", 
       subtitle = "San Joaquin Valley, June - September, 2018 - 2020", 
       caption = "") + 
  xlab("Modeled Air Temperature (C)") + 
  ylab("Observed Air Temperature (C)") + 
  ylim(c(0, 45)) + 
  xlim(c(0, 45)) + 
  geom_smooth(method = "lm", se = FALSE, aes(color = "Linear regression")) +
  guides(alpha = FALSE) +
  theme(legend.title = element_blank())
```

What are our confidence intervals for this? How much error can we expect? 
```{r, include=FALSE, echo = FALSE}
summary(final)
```

Even though our Adjusted $R^2$ is 0.8474, the residual standard error is 2.886, which is rather large. That means that we have a plus or minus 2.886*1.96=5.65656 confidence interval, which is not great. Our confidence interval is ranging over 10 C... 